---
title: "Code Challenge: Sentiment Analysis"
author: "Yongjun Zhang"
date: ""
output:
  rmdformats::readthedown:
    highlight: pygments
--- = =
---

```{=html}
<style type="text/css">
p{ /* Normal  */
   font-size: 18px;
}
body{ /* Normal  */
   font-size: 18px;
}
td {  /* Table  */
   font-size: 14px;
}
h1 { /* Header 1 */
 font-size: 32px;
}
h2 { /* Header 2 */
 font-size: 26px;
}
h3 { /* Header 3 */
 font-size: 22px;
}
code.r{ /* Code block */
  font-size: 14px;
}
pre { /* Code block */
  font-size: 14px
}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instruction

> You are required to conduct some sentiment analysis for IMDB movie reviews

> The training dataset can be loaded using read_csv(url("https://yongjunzhang/files/css/imdb_sample.csv")). 

> It is a random sample of IMDB movie review data. It has 1500 negative reviews and 1500 positive reviews.

> You are required to compute sentiment scores using one of lexicon-based methods and one of conventional ML methods.

> You need to turn in your rmarkdown file and the associated pdf or html.

> Some clues: You can use one of these r packages we mentioned in our lecture and just do mutate to create a sentiment score. For ML method, you can do caret style training or use quanteda package.

> You have two weeks to complete this challenge.

> Note that you should also work on your final research project as well.



```{r}
require(pacman)
p_load(tidyverse,glue,sentimentr,syuzhet,quanteda,quanteda.textmodels,caret)
```

```{r}
data <- read_csv(url("https://yongjunzhang.com/files/css/imdb_sample.csv"))
```

```{r}

corp_movies <- corpus(data)
summary(corp_movies, 10)

set.seed(300)
id_train <- sample(1:3000, 2500, replace = FALSE)
head(id_train, 10)
# create docvar with ID
corp_movies$id_numeric <- 1:ndoc(corp_movies)

# tokenize texts
toks_movies <- tokens(corp_movies, remove_punct = TRUE, remove_number = TRUE) %>% 
               tokens_remove(pattern = stopwords("en")) %>% 
               tokens_wordstem()
dfmt_movie <- dfm(toks_movies)

# get training set
dfmat_training <- dfm_subset(dfmt_movie, id_numeric %in% id_train)

# get test set (documents not in id_train)
dfmat_test <- dfm_subset(dfmt_movie, !id_numeric %in% id_train)

# Next we train the naive Bayes classifier using textmodel_nb().

tmod_nb <- textmodel_nb(dfmat_training, dfmat_training$sentiment)
summary(tmod_nb)
```

```{r}
# Naive Bayes can only take features into consideration that occur both in the training set and the test set, but we can make the features identical using dfm_match()

dfmat_matched <- dfm_match(dfmat_test, features = featnames(dfmat_training))

# Let’s inspect how well the classification worked.

actual_class <- dfmat_matched$sentiment
predicted_class <- predict(tmod_nb, newdata = dfmat_matched)
tab_class <- table(actual_class, predicted_class)
tab_class
```

```{r}
# We can use the function confusionMatrix() from the caret package to assess the performance of the classification.

confusionMatrix(tab_class, mode = "everything")
```


