---
title: "Code Challenge in ML: Replicating Baby Names Gender Prediction"
author: "Yongjun Zhang, Ph.D."
institute: "Department of Sociology and IACS, Stony Brook Unversity"
date: ""
output: pdf_document
--- 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data Preparation

> You can download ssa baby names via here <https://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-national-data>

> I have also used the following codes to preprocess the data. I have generated aggregated name counts by gender and then define our target outcome as "female if over 50 of women use that name." I have also created four variables: first letter, first two letters, last letter, and laster two letters. You can directly use the RData file, <https://yongjunzhang.com/files/ssa_baby_names.RData>. 

```{r}
require(pacman)
p_load(tidyverse,glue)

files <- list.files(path = "./names/",pattern = ".txt",full.names = TRUE)
files

# let us say we want to define a read_us_baby function
readSsaBabyNames <- function(file,...){
  require(tidyverse)
  data <- read_csv(file,col_names = FALSE,show_col_types = FALSE) %>% 
    mutate(year=str_replace_all(file,"[^0-9]",""))
  colnames(data) <- c("names","sex","count","year")
  return(data)
}

dat_year <- map_dfr(files, readSsaBabyNames)

# we only use names after 1970 and used by at least 10
dat_all <- dat_year %>%
  filter(year>1970) %>% 
  group_by(names,sex) %>% 
  summarise(count=sum(count)) %>% 
  # we only keep names used by at least 10 
  filter(count>10) %>% 
  pivot_wider(names_from = "sex",values_from="count") %>% 
  replace_na(list(F=0,M=0)) %>% 
  mutate(female=(F/(F+M)>.5)*1) %>% 
  mutate(flt1= str_extract(names,"^.")%>% tolower,
         flt2= str_extract(names,"^.{2}") %>% tolower,
         llt1= str_extract(names,".$")%>% tolower,
         llt2= str_extract(names,".{2}$")%>% tolower
         )
save(dat_year,dat_all,file="./ssa_baby_names.RData")
```

### Split our data into train and test data

Before we further split our data, let us take a look at data first. Since the code challenge asks us to select top 5 most frequent features, we take a look and see which letters are most frequent in the database.

```{r}

top_letters <- dat_all %>% 
  group_by(flt1) %>% 
  summarise(fl1=n()) %>% 
  top_n(n=5) %>% 
  bind_cols(
    dat_all %>% 
      group_by(flt2) %>% 
      summarise(fl2=n()) %>% 
      top_n(n=5)
  ) %>% 
  bind_cols(
    dat_all %>% 
      group_by(llt1) %>% 
      summarise(ll1=n()) %>% 
      top_n(n=5)
  ) %>% 
  bind_cols(
    dat_all %>% 
      group_by(llt2) %>% 
      summarise(ll2=n()) %>% 
      top_n(n=5)
  )

knitr::kable(top_letters)

```

### Let us create these features

including, flt1.a, flt1.j, flt1.k, flt1.m, flt1.s,flt2.da, flt2.ja, flt2.ka,flt2.ma, flt2.sh,
ll1.a, ll1.e, ll1.h, ll1.i, ll1.n, etc..


```{r}
library(caret)
#create the full datasets with dummies
top_features <- c(paste0("flt1",c("a","j","k","m","s")),
                  paste0("flt2",c("da","ja","ka","ma","sh")),
                  paste0("llt1",c("a","e","h","i","n")),
                  paste0("llt2",c("ah","an","ia","na","on"))
                  )
fllt_d=predict(dummyVars(~llt1+llt2+flt1+flt2,data=dat_all),newdata=dat_all)%>%as.data.frame()

df=dat_all %>% 
  ungroup %>% 
  select(names,female) %>% 
  filter(!is.na(female)) %>% 
  mutate(female=ifelse(female==1,"Y","N") %>% as.factor()) %>% 
  bind_cols(fllt_d %>% 
              select(all_of(top_features)))
```

Let us finally split our data into train and test

```{r}
inTrain <- createDataPartition(
  y = df$female,
  ## the outcome data are needed
  p = .75,
  ## The percentage of data in the
  ## training set
  list = FALSE
)

train <- df[ inTrain,]
test  <- df[-inTrain,]

nrow(train)
#> [1] 157
nrow(test)
library(gdata)
keep(dat_all,df,train,test,sure=TRUE)

```

### Let us train a nb model to predict gender

You need naivebayes package to be installed, here is the documentation:
<https://cran.r-project.org/web/packages/naivebayes/naivebayes.pdf>


```{r}
# K folds cross validation
# try parallel computing
library(doParallel)
library(naivebayes)
cl <- makePSOCKcluster(3)
registerDoParallel(cl)

# Define tuning grid 
grid_nb <- expand.grid(usekernel = c(TRUE, FALSE),
                         laplace = c(0, 0.5, 1), 
                         adjust = c(0.75, 1, 1.25, 1.5))

train_control <- caret::trainControl(
  method = "cv",
  number = 3,
  classProbs=T,savePredictions = T,
  verboseIter = FALSE, 
  allowParallel = TRUE 
)

nb_base <- caret::train(
  female~., 
  data=train %>%  dplyr::select(-c(names)),
  trControl = train_control,
  tuneGrid = grid_nb,
  method = "naive_bayes",
  verbose = TRUE
)

stopCluster(cl)

save(nb_base,file="./nb_base.RData")

```

### Let us check model performance, get the comfusion matrix firrst

```{r}
#load("./nb_base.RData")
# check cf matrix
nb_base
confusionMatrix(nb_base)
# PLOT ROC CURVE
library(MLeval)
## run MLeval
res <- evalm(nb_base)
## get ROC
res$roc
```

### Test model performance on test set

```{r}
# predict test names
test_df <- test %>% dplyr::select(-c(names,female))
test_pred <- predict(nb_base$finalModel,newdata=test_df) %>% as.data.frame()
colnames(test_pred) <- "nb_female"
# check confusion matrix
confusionMatrix(test_pred$nb_female,test$female)
```

### Let us try penalized logit models
we will use glmnet
-combination of lasso and ridge regression
-Can fit a mix of the two models
-alpha [0, 1]: pure lasso to pure ridge
-lambda (0, infinity): size of the penalty

```{r}
# K folds cross validation
# try parallel computing
library(glmnet)
cl <- makePSOCKcluster(3)
registerDoParallel(cl)

# Define tuning grid 
grid_plr <- expand.grid( alpha = c(0,1),
                         lambda = c(1e-4, 1e-2, 1))

train_control <- caret::trainControl(
  method = "cv",
  number = 3,
  classProbs=T,savePredictions = T,
  verboseIter = FALSE, 
  allowParallel = TRUE 
)

plr_base <- caret::train(
  female~., 
  data=train %>%dplyr::select(-c(names)),
  trControl = train_control,
  tuneGrid = grid_plr,
  method = "glmnet",
  verbose = TRUE
)

stopCluster(cl)
save(plr_base,file="./plr_base.RData")
# check cf matrix
plr_base
confusionMatrix(plr_base)
```


# Let us try random forest

Random Forest

-method = 'ranger'
-Type: Classification, Regression

Tuning parameters:

-mtry (#Randomly Selected Predictors)
-splitrule (Splitting Rule)
-min.node.size (Minimal Node Size)
-Required packages: e1071, ranger, dplyr

here is the documentation: <https://cran.r-project.org/web/packages/ranger/ranger.pdf>

```{r}
# K folds cross validation
# try parallel computing
require(pacman)
p_load(ranger,e1071)
cl <- makePSOCKcluster(3)
registerDoParallel(cl)

# Define tuning grid 
grid_rf <- expand.grid(splitrule= c("gini", "extratrees","hellinger"),
                      mtry = c(1,2,4,10),
                      min.node.size=c(1))

train_control <- caret::trainControl(
  method = "cv",
  number = 3,
  classProbs=T,savePredictions = T,
  verboseIter = FALSE, 
  allowParallel = TRUE 
)

rf_base <- caret::train(
  female~., 
  data=train %>%dplyr::select(-c(names)),
  trControl = train_control,
  tuneGrid = grid_rf,
  method ='ranger',
  verbose = TRUE
)

stopCluster(cl)
save(rf_base,file="./rf_base.RData")
# check cf matrix
rf_base
confusionMatrix(rf_base)
```


